---
title: "Data-analysis"
author: "Zeinab Al Ahmad"
date: "2024-02-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
## Welcome to the Pharmacological 2024 Data!
Here I practiced some charts from the 2024 Pharma data since I don't yet have data for myself!


```{r message=FALSE, warning=FALSE}
#Libraries 
library(readxl)
library(tidyverse) 
library(readxl) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(igraph)

```


```{r}
# Read Ligand data from CSV files
Ligands <-read.csv("C:/Users/zalah/OneDrive/Documents/Wflowr.Zee/data/ligands.csv")
```

## 12590 Ligands Classfied by their types 
Different types of ligands: Antibody, Inorganic, Metabolite, Natural Product, Peptide, and SYnthetic Organic 
This graph is a bar chart that illustrates the classification of ligands by type.
The x-axis represents different ligand types, namely "Antibody," "Inorganic," "Metabolite," "Natural product," "Peptide," and "Synthetic organic." 
The y-axis indicates the count of each ligand type, ranging from 0 to 9000. 
The bars' height represents the frequency of each ligand type, with "Synthetic organic" being the most common ligand in this dataset.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Create a data frame that summarizes ligand names by type, along with the count
ligands_summary <- Ligands %>%
  group_by(Type) %>%
  summarise(
    Count = n(),  # Count the number of ligands per type
    Names = paste(Name, collapse = "; ")  
  ) %>%
  ungroup()  

# Create the interactive plot using plotly
interactive_plot <- plot_ly(
  data = ligands_summary,
  x = ~Type,
  y = ~Count,
  type = 'bar',
  text = ~Names,
  hoverinfo = 'text',  # Display names on hover
  marker = list(color = 'pink')
)

# Customize the layout
interactive_plot <- interactive_plot %>%
  layout(
    title = "Ligands Classification by Type",
    xaxis = list(title = "Type of Ligand"),
    yaxis = list(title = "Count"),
    hoverlabel = list(align = "left")  # Align hover labels to the left
  )

# Render the plot
interactive_plot
```



```{r echo=TRUE, message=FALSE, warning=FALSE}

Ligands$Approved <- as.factor(Ligands$Approved)


```

## Heatmap for Ligands Classification
Ligands Classified by Type and Approval Status
(Not all ligand names were shown on the Y-axis but if you use your cursor over the map you'll see all names and if approved or not z=0 not approved/ z=1 approved)

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(plotly)


interactive_plot <- plot_ly(data = Ligands, x = ~Type, y = ~Name, text = ~Approved,
                            mode = 'markers', color = ~Approved, colors = c('red', 'green')) %>%
  layout(title = 'Ligands by Type and Approval Status',
         xaxis = list(title = 'Type of Ligand'),
         yaxis = list(categoryorder = 'total ascending', title = 'Ligand Name'),
         hovermode = 'closest')
interactive_plot
```

## Easier way to look?
Table Below shows 12,590 Ligands' Names, Species, Type, and if Approved or Not
You can click the search bar and search for any Ligand Name.

```{r echo=TRUE, message=FALSE, warning=FALSE}

library(DT)
datatable(Ligands[, c("Ligand.ID", "Name", "Species", "Type", "Approved")], 
          filter = 'top', 
          options = list(pageLength = 10, scrollX = TRUE))

 
```

## Another Ligands Classification 
Ligands classified according to Ligands type and approval status 

```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(Ligands, aes(x = Approved)) +
  geom_bar(fill = "pink") +
  facet_wrap(~Type, scales = "free_x") +
  theme_minimal() +
  labs(title = "Number of Ligands by Type and Approval Status", 
       x = "Approval Status", 
       y = "Count of Ligands") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r echo=TRUE, message=FALSE, warning=FALSE}
approval_summary <- Ligands %>%
  count(Approved)

print(approval_summary)
```


## These Ligands were then tested on different Targets

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Read Interactions data from CSV files (I skipped first row since it is empty)
interactions <-read.csv("C:/Users/zalah/OneDrive/Documents/Wflowr.Zee/data/interactions.csv", skip = 1)
# Select the desired columns
Filtered_interactions <- interactions %>% select(Target, Ligand, Ligand.Type, Approved, Action, Target.Species)
# Print the first few rows to check the result
head(Filtered_interactions)
```

## Classification of Ligands by their action
each drug has a different kind of action on a specific target as shown in the graph below

```{r}
# Group by Action and summarize the count of Ligands
action_summary <- Filtered_interactions %>%
  group_by(Action) %>%
  summarize(Ligand_Count = n()) %>%
  arrange(desc(Ligand_Count))

# Plot the data using ggplot2
ggplot(action_summary, aes(x = reorder(Action, -Ligand_Count), y = Ligand_Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Ligands Classification by Action",
       x = "Action",
       y = "Count of Ligands") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```






## Relationship between Ligand and Target by Approval Status
first 25 rows only

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Select the first 20 rows and the necessary columns: Ligand, Target, and Approved
data_for_plot <- interactions %>%
  select(Ligand, Target, Approved) %>%
  slice(1:25) %>%
  distinct()

ggplot(data_for_plot, aes(x = Target, y = Ligand, color = Approved)) +
  geom_point(position = position_jitter(width = 0.2, height = 0), size = 1) +
  labs(title = "Relationship between Ligand and Target by Approval Status",
       x = "Target",
       y = "Ligand") +
  scale_color_manual(values = c("red", "blue"), labels = c("Not Approved", "Approved")) +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 0, hjust = 0, size = 7),  # Rotate y-axis labels and adjust size
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = element_blank())

```


## Relationship between Ligand and Target by Approval status using Plotly
For names not to overlapp i removed the names of Ligands and Targets from the X-axis and Y-axis and kept it like this (if you press on any dot you'll see the name of ligand, target and if approved or not!) That way I can do more than just 25 rows (5000 rows presented here)


```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(dplyr)


data_for_plot <- interactions %>%
  select(Ligand, Target, Approved) %>%
  slice(1:5000) %>%
  distinct()

# Create a ggplot
p <- ggplot(data_for_plot, aes(x = Target, y = Ligand, color = Approved)) +
  geom_point(position = position_jitter(width = 0.2, height = 0), size = 0.5) +
  labs(title = "Relationship between Ligand and Target by Approval Status",
       x = "Target",  # Remove x-axis label
       y = "Ligand") +  # Remove y-axis label
  scale_color_manual(values = c("red", "blue"), labels = c("Not Approved", "Approved")) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),  # Remove x-axis text
        axis.text.y = element_blank(),  # Remove y-axis text
        axis.ticks = element_blank(),  # Remove axis ticks
        legend.title = element_blank())

# Save the plot
ggsave("plot.png", plot = p, width = 15, height = 10, units = "in")

# Convert to interactive plot
ggplotly(p)


```









